---
output: 
    github_document:
        toc: true
        toc_depth: 2
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# About `coefixr`

<!-- badges: start -->
<!-- badges: end -->

`coefixr` adjusts the coefficients and confidence intervals associated with interaction terms in models, including interactions that involve the reference levels of Factors. It should work with a variety of models.


# Installation

You can install the development version of `coefixr` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("DesiQuintans/coefixr")
```

It is not currently on CRAN, nor are there plans to release it there.


# Functions available

| Function                     | Description                                                                                         |
| :--------------------------- | :-------------------------------------------------------------------------------------------------- |
| `adjust_interaction_model()` | Takes a model and returns a dataframe of adjusted coefficients, confidence intervals, and p-values. |
| `cancer_modified`            | A built-in dataset for testing.                                                                     |



# Worked example

Here is an example of fitting a linear model with an interaction on a dataset that is included with `coefixr`.

```{r}
library(coefixr)

# cancer_modified is a dataset provided with the `coefixr` package.
str(cancer_modified)

Map(levels, Filter(is.factor, cancer_modified))
```

This is the example model. It has two interactions: one between two Factors (`sex * ph.ecog`) and another between a Factor and a Numeric (`sex * wt.loss`).

```{r, echo = FALSE}
options(width = 999)
```

```{r}
my_model <- 
    lm(status ~ inst + age + sex * ph.ecog + sex * wt.loss, 
       data = cancer_modified)

summary(my_model)$coefficients
```


### Automated calculation with `adjust_interaction_model()`

We can use `adjust_interaction_model()` to adjust the coefficients and confidence intervals of the interactions. For details of those adjustments, see the [Calculations section](#calculations). 

Here, I request non-exponentiated coefficients, a global Type-II p-value for each variable, and output formatted to two decimal places for all numbers. If I wanted unrounded numbers, I would set `digits.n` and `digits.p` to `Inf`.

```{r}
adjust_interaction_model(
    modelobj     = my_model, 
    data         = cancer_modified, 
    exponentiate = FALSE,
    add.global.p = TRUE,
    digits.n     = 2, 
    digits.p     = 2,
    global_args  = list(type = "II")
)
```

`adjust_interaction_model()` returns a data frame with these columns:

- **covar**
    - The covariate.
- **ref**
    - `TRUE` marks the reference levels of covariates.
- **ref.intx**
    - `TRUE` marks interactions that involve the reference level of a covariate.
- **global.p**
    - The global p-value of the covariate. Column is omitted if `add.global.p = FALSE`.
- **p.value**
    - The p-value of the covariate.
- **log_ci.95lwr** or **exp_ci.95lwr**
    - Lower 95% confidence interval of the coefficient. If `exponentiate = TRUE`, the column's name is changed and the contents are exponentiated.
- **log_coef** or **exp_coef**
    - The coefficient. If `exponentiate = TRUE`, the column's name is changed and the contents are exponentiated.
- **log_ci.95upr** or **exp_ci.95upr**
    - Upper 95% confidence interval of the coefficient. If `exponentiate = TRUE`, the column's name is changed and the contents are exponentiated.



# Calculations

## Adjusted interaction coefficients

```{r}
summary(my_model)$coefficients
```

Coefficients involved in interactions are adjusted by adding the coefficient of the interaction and all of its components together. For example, the interaction of `sexMale:ph.ecogNot completely ambulatory` involves:

- `sexMale` ($0.471133734$)
- `ph.ecogNot completely ambulatory` ($0.465297213$)
- `sexMale:ph.ecogNot completely ambulatory` ($-0.289047417$)

$$0.471133734 + 0.465297213 + -0.289047417 = 0.6473835$$


## Adjusted standard error (SE)

```{r}
covm <- vcov(my_model)
covm[upper.tri(covm, diag = FALSE)] <- NA

covm
```


The standard error (SE) is adjusted using $\sqrt{\text{variances} + 2(\text{covariances})}$ for all terms involved in the interaction. The variance-covariance table is accessed using `stats::vcov()`. For example, using `vcov(my_model)`, the interaction of `sexMale:ph.ecogNot completely ambulatory` involves:

- The variances
    - `sexMale` ($0.01303826$)
    - `ph.ecogNot completely ambulatory` ($0.01802702$)
    - `sexMale:ph.ecogNot completely ambulatory` ($0.03024963$)
- The covariances
    - `sexMale` vs `ph.ecogNot completely ambulatory` ($0.007501756$)
    - `sexMale` vs `sexMale:ph.ecogNot completely ambulatory` ($-0.01191112$)
    - `ph.ecogNot completely ambulatory` vs `sexMale:ph.ecogNot completely ambulatory` ($-0.01775626$)

$$\sqrt{0.01303826 + 0.01802702 + 0.03024963 + 2(0.007501756 + -0.01191112 + -0.01775626)} = 0.1303214$$


## Adjusted 95% confidence interval

The adjusted confidence interval is calculated by subtracting/adding $1.96 \times SE$ from the adjusted coefficient. Again, for the interaction of `sexMale:ph.ecogNot completely ambulatory` which has an adjusted coefficient of $0.6473835$ and an adjusted SE of $0.1303214$:

$$\text{Lower 95% CI:} 0.6473835 - 19.6 \times 0.1303214 = 0.3919536$$
$$\text{Upper 95% CI:} 0.6473835 + 19.6 \times 0.1303214 = 0.9028134$$
