---
output: 
    github_document:
        toc: true
        toc_depth: 2
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- PDF manual is created in terminal with:  R CMD Rd2pdf . -o coefixr.pdf -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```



# About `coefixr`

<!-- badges: start -->
<!-- badges: end -->

`coefixr` adjusts the coefficients and confidence intervals associated with interaction terms in models, including interactions that involve the reference levels of Factors. It should work with a variety of models.


# Installation

You can install the development version of `coefixr` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("DesiQuintans/coefixr")
```

It is not currently on CRAN, nor are there plans to release it there.


# Package contents

| Function                     | Description                                                                              |
| :--------------------------- | :--------------------------------------------------------------------------------------- |
| `adjust_interaction_model()` | Takes a model and returns a dataframe of adjusted coefficients and confidence intervals. |
| `cancer_modified`            | A built-in dataset for testing.                                                          |


# Package manual

<https://github.com/DesiQuintans/coefixr/blob/main/coefixr.pdf>


# Worked example

Here is an example of fitting a linear model with an interaction on a dataset that is included with `coefixr`.

```{r}
library(coefixr)

# cancer_modified is a dataset provided with the `coefixr` package.
str(cancer_modified)

Map(levels, Filter(is.factor, cancer_modified))
```

This is the example model. It has two interactions: one between two Factors (`sex * ph.ecog`) and another between a Factor and a Numeric (`sex * wt.loss`).

```{r, echo = FALSE}
options(width = 999)
```

```{r}
my_model <- 
    lm(status ~ inst + age + sex * ph.ecog + sex * wt.loss, 
       data = cancer_modified)

summary(my_model)$coefficients
```


### Automated calculation with `adjust_interaction_model()`

We can use `adjust_interaction_model()` to adjust the coefficients and confidence intervals of the interactions. For details of those adjustments, see the [Calculations section](#details-re-the-calculations-performed). 

Here, I request non-exponentiated (i.e. log) coefficients, a global Type-II p-value for each variable, and output formatted to 2 decimal places for coefficients and 3 decimals places for p-values.

```{r}
adjust_interaction_model(
    modelobj     = my_model, 
    data         = cancer_modified, 
    exponentiate = FALSE,
    add.global.p = TRUE,
    intercept    = TRUE,
    digits.n     = 2, 
    digits.p     = 3,
    global_args  = list(type = "II")
)
```

`adjust_interaction_model()` returns a data frame with these columns:

- **covar**
    - The covariate.
- **adj.with**
    - Terms included in the adjustment of interactions. `Unadjusted` for non-interactions. See [Advanced Usage](#variables-of-interest) for more details.
- **is.top**
    - `TRUE` marks top-level covariates (i.e. names used in the model formula).
- **is.intx**
    - `TRUE` marks all interactions.
- **ref**
    - `TRUE` marks the reference levels of covariates.
- **ref.intx**
    - `TRUE` marks interactions that involve the reference level of a covariate.
- **global.p**
    - The global p-value of the covariate. Column is omitted if `add.global.p = FALSE`.
- **p.value**
    - The p-value of the covariate. P-values are inherited from the model's `summary()` method and not recalculated.
- **ci.95lwr** or **exp_ci.95lwr**
    - Lower 95% confidence interval of the coefficient. If `exponentiate = TRUE`, the column's name is changed and the contents are exponentiated.
- **coef** or **exp_coef**
    - The coefficient. If `exponentiate = TRUE`, the column's name is changed and the contents are exponentiated.
- **ci.95upr** or **exp_ci.95upr**
    - Upper 95% confidence interval of the coefficient. If `exponentiate = TRUE`, the column's name is changed and the contents are exponentiated.



# Details re. the calculations performed

## Adjusted interaction coefficients

```{r}
summary(my_model)$coefficients
```

Coefficients involved in interactions are adjusted by adding the coefficient of the interaction and all of its components together. Reference levels of Factors are set to 0.

For example, the interaction of `sexMale:ph.ecogNot completely ambulatory` involves:

- `sexMale` ($0.471133734$)
- `ph.ecogNot completely ambulatory` ($0.465297213$)
- `sexMale:ph.ecogNot completely ambulatory` ($-0.289047417$)

$$0.471133734 + 0.465297213 + -0.289047417 = 0.6473835$$


## Adjusted standard error (SE)

```{r}
covm <- vcov(my_model)
covm[upper.tri(covm, diag = FALSE)] <- NA

covm
```


The standard error (SE) is adjusted using $\sqrt{\sum\text{variances} + 2(\sum\text{covariances})}$ for all terms involved in the interaction. The variance-covariance table is accessed using `stats::vcov()`. Terms that do not appear in the `vcov()` matrix (namely, reference levels and interactions involving reference levels) are set to 0. For example, using `vcov(my_model)`, the interaction of `sexMale:ph.ecogNot completely ambulatory` involves:

- The variances
    - `sexMale` ($0.01303826$)
    - `ph.ecogNot completely ambulatory` ($0.01802702$)
    - `sexMale:ph.ecogNot completely ambulatory` ($0.03024963$)
- The covariances
    - `sexMale` vs `ph.ecogNot completely ambulatory` ($0.007501756$)
    - `sexMale` vs `sexMale:ph.ecogNot completely ambulatory` ($-0.01191112$)
    - `ph.ecogNot completely ambulatory` vs `sexMale:ph.ecogNot completely ambulatory` ($-0.01775626$)

$$\sqrt{0.01303826 + 0.01802702 + 0.03024963 + 2(0.007501756 + -0.01191112 + -0.01775626)} = 0.1303214$$


## Adjusted 95% confidence interval

The adjusted confidence interval is calculated by subtracting/adding $1.96 \times SE$ from the adjusted coefficient. Again, for the interaction of `sexMale:ph.ecogNot completely ambulatory` which has an adjusted coefficient of $0.6473835$ and an adjusted SE of $0.1303214$:

$$\text{Lower CI} = 0.6473835 - 1.96 \times 0.1303214 = 0.3919536$$
$$\text{Upper CI} = 0.6473835 + 1.96 \times 0.1303214 = 0.9028134$$



## Comparison to package output

```{r}
res <- adjust_interaction_model(my_model, cancer_modified, digits.n = Inf)

res[which(res$covar == "sexMale:ph.ecogNot completely ambulatory"),
    c("covar", "ci.95lwr", "coef", "ci.95upr")]
```


# Advanced usage

## Variables of interest

If you want the coefficients, SEs, and confidence intervals of your interactions to be adjusted only using some variable(s) of interest, you can specify this through the `interest` argument. 

For example, adjustment of the interaction of `sexMale:ph.ecogNot completely ambulatory` in the worked example uses the coefficients/variances/covariances of these three terms:

- `sexMale`
- `ph.ecogNot completely ambulatory`
- `sexMale:ph.ecogNot completely ambulatory`

If you only want to look at the effect of sex in the interaction, then running:

```{r eval=FALSE, include=TRUE}
adjust_interaction_model(
    modelobj = my_model, 
    data     = cancer_modified, 
    interest = "sex"
)
```

would perform the adjustment using any term that matches the regular expression `sex` and ignoring all others. Which is to say, it would use 

- `sexMale`
- `sexMale:ph.ecogNot completely ambulatory`

And not use the fixed effect of `ph.ecogNot completely ambulatory` because it does not match. Here it is in action:

## Worked example of adjusting with `interest = "sex"`

```{r eval=FALSE, include=TRUE}
my_model <- 
    lm(status ~ inst + age + sex * ph.ecog + sex * wt.loss, 
       data = cancer_modified)

adjust_interaction_model(
    modelobj = my_model, 
    data     = cancer_modified, 
    interest = "sex"
)
```


### Adjusted coefficients

- `sexMale` ($0.471133734$)
- (IGNORED) `ph.ecogNot completely ambulatory`
- `sexMale:ph.ecogNot completely ambulatory` ($-0.289047417$)

$$0.471133734 + -0.289047417 = 0.1820863$$

### Adjusted SE

- The variances
    - `sexMale` ($0.01303826$)
    - (IGNORED) `ph.ecogNot completely ambulatory`
    - `sexMale:ph.ecogNot completely ambulatory` ($0.03024963$)
- The covariances
    - (IGNORED) `sexMale` vs `ph.ecogNot completely ambulatory`
    - `sexMale` vs `sexMale:ph.ecogNot completely ambulatory` ($-0.01191112$)
    - (IGNORED) `ph.ecogNot completely ambulatory` vs `sexMale:ph.ecogNot completely ambulatory`

$$\sqrt{0.01303826 + 0.03024963 + 2(-0.01191112)} = 0.139519381$$

### Adjusted CI

$$\text{Lower CI} = 0.1820863 - 1.96 \times 0.139519381 = -0.09137169$$
$$\text{Upper CI} = 0.1820863 + 1.96 \times 0.139519381 = 0.4555443$$


### Comparison to package output

```{r}
res <- adjust_interaction_model(my_model, cancer_modified, interest = "sex")

res[which(res$covar == "sexMale:ph.ecogNot completely ambulatory"),
    c("covar", "ci.95lwr", "coef", "ci.95upr")]
```


### Nota bene

Because `interest` is a case-sensitive regular expression, you have a lot of flexibility about what you want to specify, like telling it that you'd like to adjust by any term that first matches "ecog" and then later matches "ymp", which means that you would be adjusting with `ph.ecogAsymptomatic` or `ph.ecogSymptomatic but completely ambulatory`, but NOT `ph.ecogNot completely ambulatory`:

```{r eval=FALSE, include=TRUE}
adjust_interaction_model(
    modelobj = my_model, 
    data     = cancer_modified, 
    interest = "ecog.*?ymp"
)
```

It's left to the user to know what they are asking for and why.
