---
output: 
    github_document:
        toc: true
        toc_depth: 2
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# About `coefixr`

<!-- badges: start -->
<!-- badges: end -->

`coefixr` adjusts the coefficients and confidence intervals associated with interaction terms in models, including interactions that involve the reference levels of Factors. It should work with a variety of models.


# Installation

You can install the development version of `coefixr` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("DesiQuintans/coefixr")
```

It is not currently on CRAN, nor are there plans to release it there.


# Package contents

| Function                     | Description                                                                              |
| :--------------------------- | :--------------------------------------------------------------------------------------- |
| `adjust_interaction_model()` | Takes a model and returns a dataframe of adjusted coefficients and confidence intervals. |
| `cancer_modified`            | A built-in dataset for testing.                                                          |



# Worked example

Here is an example of fitting a linear model with an interaction on a dataset that is included with `coefixr`.

```{r}
library(coefixr)

# cancer_modified is a dataset provided with the `coefixr` package.
str(cancer_modified)

Map(levels, Filter(is.factor, cancer_modified))
```

This is the example model. It has two interactions: one between two Factors (`sex * ph.ecog`) and another between a Factor and a Numeric (`sex * wt.loss`).

```{r, echo = FALSE}
options(width = 999)
```

```{r}
my_model <- 
    lm(status ~ inst + age + sex * ph.ecog + sex * wt.loss, 
       data = cancer_modified)

summary(my_model)$coefficients
```


### Automated calculation with `adjust_interaction_model()`

We can use `adjust_interaction_model()` to adjust the coefficients and confidence intervals of the interactions. For details of those adjustments, see the [Calculations section](#details-re-the-calculations-performed). 

Here, I request non-exponentiated (i.e. log) coefficients, a global Type-II p-value for each variable, and output formatted to 2 decimal places for coefficients and 3 decimals places for p-values.

```{r}
adjust_interaction_model(
    modelobj     = my_model, 
    data         = cancer_modified, 
    exponentiate = FALSE,
    add.global.p = TRUE,
    intercept    = TRUE,
    digits.n     = 2, 
    digits.p     = 3,
    global_args  = list(type = "II")
)
```

`adjust_interaction_model()` returns a data frame with these columns:

- **covar**
    - The covariate.
- **is.top**
    - `TRUE` marks top-level covariates (i.e. names used in the model formula).
- **is.intx**
    - `TRUE` marks all interactions.
- **ref**
    - `TRUE` marks the reference levels of covariates.
- **ref.intx**
    - `TRUE` marks interactions that involve the reference level of a covariate.
- **global.p**
    - The global p-value of the covariate. Column is omitted if `add.global.p = FALSE`.
- **p.value**
    - The p-value of the covariate. P-values are inherited from the model's `summary()` method and not recalculated.
- **ci.95lwr** or **exp_ci.95lwr**
    - Lower 95% confidence interval of the coefficient. If `exponentiate = TRUE`, the column's name is changed and the contents are exponentiated.
- **coef** or **exp_coef**
    - The coefficient. If `exponentiate = TRUE`, the column's name is changed and the contents are exponentiated.
- **ci.95upr** or **exp_ci.95upr**
    - Upper 95% confidence interval of the coefficient. If `exponentiate = TRUE`, the column's name is changed and the contents are exponentiated.



# Details re. the calculations performed

## Adjusted interaction coefficients

```{r}
summary(my_model)$coefficients
```

Coefficients involved in interactions are adjusted by adding the coefficient of the interaction and all of its components together. Reference levels of Factors are set to 0.

For example, the interaction of `sexMale:ph.ecogNot completely ambulatory` involves:

- `sexMale` ($0.471133734$)
- `ph.ecogNot completely ambulatory` ($0.465297213$)
- `sexMale:ph.ecogNot completely ambulatory` ($-0.289047417$)

$$0.471133734 + 0.465297213 + -0.289047417 = 0.6473835$$


## Adjusted standard error (SE)

```{r}
covm <- vcov(my_model)
covm[upper.tri(covm, diag = FALSE)] <- NA

covm
```


The standard error (SE) is adjusted using $\sqrt{\sum\text{variances} + 2(\sum\text{covariances})}$ for all terms involved in the interaction. The variance-covariance table is accessed using `stats::vcov()`. Terms that do not appear in the `vcov()` matrix (namely, reference levels and interactions involving reference levels) are set to 0. For example, using `vcov(my_model)`, the interaction of `sexMale:ph.ecogNot completely ambulatory` involves:

- The variances
    - `sexMale` ($0.01303826$)
    - `ph.ecogNot completely ambulatory` ($0.01802702$)
    - `sexMale:ph.ecogNot completely ambulatory` ($0.03024963$)
- The covariances
    - `sexMale` vs `ph.ecogNot completely ambulatory` ($0.007501756$)
    - `sexMale` vs `sexMale:ph.ecogNot completely ambulatory` ($-0.01191112$)
    - `ph.ecogNot completely ambulatory` vs `sexMale:ph.ecogNot completely ambulatory` ($-0.01775626$)

$$\sqrt{0.01303826 + 0.01802702 + 0.03024963 + 2(0.007501756 + -0.01191112 + -0.01775626)} = 0.1303214$$


## Adjusted 95% confidence interval

The adjusted confidence interval is calculated by subtracting/adding $1.96 \times SE$ from the adjusted coefficient. Again, for the interaction of `sexMale:ph.ecogNot completely ambulatory` which has an adjusted coefficient of $0.6473835$ and an adjusted SE of $0.1303214$:

$$\text{Lower CI} = 0.6473835 - 19.6 \times 0.1303214 = 0.3919536$$
$$\text{Upper CI} = 0.6473835 + 19.6 \times 0.1303214 = 0.9028134$$



## Comparison to package output

```{r}
res <- adjust_interaction_model(my_model, cancer_modified, digits.n = Inf)

res[which(res$covar == "sexMale:ph.ecogNot completely ambulatory"),
    c("covar", "ci.95lwr", "coef", "ci.95upr")]
```

